---
title: "Script 2"
author: "Silvy van Kuijk, Sun O'Brien & Anthony Di Fiore"
date: "2023-08-25"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Install Dependencies
This script was written in R version 4.1.2.

```{r, include = FALSE}
# Package names
packages <- c("tidyverse", # For data wrangling, also loads in lubridate, dplyr, ggplot2.
              "monitoR", # For automated detection, also loads in tuneR.
              "foreach",
              "brms",
              "MASS",
              "data.table",
              "caret") 
# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# Load packages
lapply(packages, require, character.only = TRUE)    # Load multiple packages
```

# Load Datasets

```{r, include = FALSE}
# Step 1 data
audioFileList <- read_csv("Data/audioFileList.csv", col_names = TRUE)
```

# Step 1.1: Create and test your templates
Create a list of all audio files that you want to use for template creation. For
simplicity, we will only use a few of our templates as examples here. This will 
mean that results of the automated detection algorithm here will not match the 
results in our paper as we use a limited set of templates and a limited set of 
audio files as test data for demonstration purposes.

```{r}
fileList <- list.files(path = "Audio/AutomatedDetection/Templates", full.names = TRUE) # Get a list of all files in directory.
```

Creating templates with the monitoR package is a fairly straightforward process. 
You can read more details on the process in Katz et al. (2016) or this vignette:
https://cran.r-project.org/web/packages/monitoR/monitoR.pdf. However, creating 
templates that capture your signals of interest well requires some trial and error. 
In creating templates, test a variation of frequency ranges, time frames and 
amplitude cutoffs to determine what works well for each template.In our experience, 
setting the upper and lower frequency limit too close to the actual frequency range 
of the call led to a lower accuracy. Templates scored better when a little ambient 
noise above and below the frequencies of interest was included. 

You'll also want to play around with combinations of templates to see what gives 
you the best results. We found that the combination of templates that worked best
included calls of different quality (calls both close to and at larger distance 
from the recorder).
Use a training dataset to test your templates. The code needed to test your data 
is similar to the code you'll run when analyzing your actual dataset.

We demonstrate some of our final template settings with the five samples below.
Templates 1, 2 and 3 are titi monkey templates, 4 and 5 are howler monkey templates.

```{r}
template1 <- makeBinTemplate("Audio/AutomatedDetection/Templates/TestSample_001.wav",
                              frq.lim = c(0.3, 1.6), t.lim = c(25, 35), 
                              name = "T1,025,-25,T", amp.cutoff = (-25))

template2 <- makeBinTemplate("Audio/AutomatedDetection/Templates/TestSample_002.wav",
                              frq.lim = c(0.2, 1.8), t.lim = c(0, 10), 
                              name = "T2,000,-25,T", amp.cutoff = (-25))

template3 <- makeBinTemplate("Audio/AutomatedDetection/Templates/TestSample_003.wav",
                              frq.lim = c(0.2, 1.8), t.lim = c(0, 10), 
                              name = "T3,005,-25,T", amp.cutoff = (-25))

template4 <- makeBinTemplate("Audio/AutomatedDetection/Templates/TestSample_012.wav",
                              frq.lim = c(0.2, 1.4), t.lim = c(3, 13), 
                              name = "T4,005,-25,H", amp.cutoff = (-25))

template5 <- makeBinTemplate("Audio/AutomatedDetection/Templates/TestSample_007.wav",
                              frq.lim = c(0.2, 1.4), t.lim = c(17, 27), 
                              name = "T5,017,-25,H", amp.cutoff = (-25))

# Create a combined template bin.
templates <- combineBinTemplates(template1,
                                 template2, 
                                 template3, 
                                 template4,
                                 template5)

```


# Step 1.2: Analyze your data using monitoR's binMatch() function and your templates

```{r}
# Create a text file to store the output.
outfile <- "detection_output.txt"
sink(file = outfile)

# Loop through all .wav files and run binMatch().
# This code generates one output file (.txt) containing "scores" for each .wav file.
for (i in 1:nrow(audioFileList)) { 
  print("------------")
  print(paste("FolderPath:", dirname(audioFileList[i,]$Path)))
  print(paste("FileName:", basename(audioFileList[i,]$filename)))
  print(paste("StartTime:", audioFileList[i,]$StartTime))
  print(paste("DistanceScore:", audioFileList[i,]$DistanceScore))
  print(paste("RecordingDate:", audioFileList[i,]$RecordingDate))
  print(paste("Species:", audioFileList[i,]$Species))
  print(paste("Detected:", audioFileList[i,]$Detected))
  scores <- binMatch(audioFileList[i,]$Path, 
                     templates, 
                     quiet = TRUE) 
  print(scores)
}

sink()
```


# Parse the "detection_output.txt" file and return a .csv file with the desired elements.
```{r}

# initialize results data frame to hold results from all audio files and templates
template.results <- tibble(
  file.path = character(),
  file.name = character(),
  recording.date = character(), 
  species = character(),
  detected = character(),
  distance.score = character(),
  template = character(),
  min.score = numeric(),
  max.score = numeric(),
  n.scores = numeric())

# initialize r dataframe to hold results from all templates for one audiofile
r <- tibble(
  file.path = character(),
  file.name = character(),
  recording.date = character(),
  species = character(),
  detected = character(),
  distance.score = character(),
  template = character(),
  min.score = numeric(),
  max.score = numeric(),
  n.scores = numeric())

# initialize vectors to fill r
file.path <- character()
recording.date <- character()
distance.score <- character()
detected <- character()
species <- character()
file.name <- character()
template <- character()
min.score <- numeric()
max.score <- numeric()
n.scores <- numeric()

f <- "detection_output.txt"
d <- read_lines(f, skip_empty_rows = TRUE) # read lines of "detection_output.txt" into a vector of lines 
fileindex <- 0 # tally of audiofiles
for (i in 1:length(d)) { # Loop through file line by line
  # first four lines clean odd and extraneous characters from the line
  d[i] <- str_remove_all(d[i],"[\"]")
  d[i] <- str_remove_all(d[i],"\\[1\\] ")
  d[i] <- str_replace_all(d[i],"  "," ")
  d[i] <- str_replace_all(d[i],": ",":")

  if (str_detect(d[i], "FolderPath:") == TRUE) {
    file.path <- str_remove(d[i], "FolderPath:") # Removes start of line
  }
  if (str_detect(d[i], "FileName:") == TRUE) {
    file.name <- str_remove(d[i], "FileName:") # Removes start of line
  }
  if (str_detect(d[i], "Detected:") == TRUE) {
    detected <- str_remove(d[i], "Detected:") # Removes start of line
  }
  if (str_detect(d[i], "RecordingDate:") == TRUE) {
    recording.date <- str_remove(d[i], "RecordingDate:") # Removes start of line
    recording.date <- parse_date_time(recording.date,"mdy", tz = "America/Bogota") %>% date() %>% as.character() # reformats recording date as a string with YYYY-MM-DD format
  }
  if (str_detect(d[i], "DistanceScore:") == TRUE) {
    distance.score <- str_remove(d[i], "DistanceScore:") # Removes start of line
  }
  if (str_detect(d[i], "Species:") == TRUE) {
    species <- str_remove(d[i], "Species:") # Removes start of line
  }
  if (str_detect(d[i], "Based on the survey file:") == TRUE) { # detects each audiofile
fileindex <- fileindex + 1 
    survey_file_name <- str_remove(d[i], "Based on the survey file:") # removes start of line
    survey_file_name <- str_remove(survey_file_name, " $") # removes terminal space
    index <- 0
  }

  l <- str_split(d[i], "[ ]+") # splits each row on spaces
  if (str_starts(l[[1]][1], "T") == TRUE) { # detects results lines for each template
    index <- index + 1
    template[index] <- l[[1]][1]
    min.score[index] <- as.numeric(l[[1]][2])
    max.score[index] <- as.numeric(l[[1]][3])
    n.scores[index] <- as.numeric(l[[1]][4])
  }

  if (i > 1 & (str_detect(l[[1]][1], "------------") == TRUE | i == length(d))) { # detects start of a block of new audiofile results or the end of file...
    # and creates a tibble for each set of results
    r <- tibble(
      # based on recycling of vectors; file.path, recording.date, distance.score,             file.name, and species are repeated for each template
      file.path = file.path,
      recording.date = recording.date,
      distance.score = distance.score,
      file.name = file.name,
      species = species,
      detected = detected,
      # the template and stats are already vectors
      template = template,
      min.score = min.score,
      max.score = max.score,
      n.scores = n.scores
    )
    template.results <- bind_rows(template.results, r) # bundle all results together
  }
}

# cleanup memory and work space
rm(list = c(
  "d",
  "l",
  "f",
  "r",
  "fileindex",
  "index",
  "survey_file_name",
  "file.path",
  "species",
  "recording.date",
  "file.name",
  "detected",
  "distance.score",
  "template",
  "min.score",
  "max.score",
  "n.scores",
  "i"
))
```


# Step 1.3: Average results of all templates for each audio file. 
Next we calculate the average of the template detection results. For each audio file, the binMatch() function reports a max score and a min score. The max score is of interest to us, as it gives us the highest similarity score for that audio file. We average the max scores of all templates for each audio file.

```{r}
# Split current 'template' variable into two: 'template' (template name, identical to templates names used in lines 74-102 and 'template.species' (titi (T) or howler (H) template).
template.summary <- template.results %>% mutate(
    template = str_sub(template.results$template, 1, 2),
    template.species = str_sub(template.results$template, -1, -1)
  )

# Calculate average score for both howler and titi templates.
template.summary <- template.summary %>% group_by(file.path, file.name, recording.date, species, detected, distance.score, template.species ) %>%
  summarize(mean.detection.score = mean(max.score))

# Pivot table orientation
template.summary <- template.summary %>% 
  pivot_wider(., names_from=c("template.species"),
              values_from = c("mean.detection.score"))

```

Next, determine which scores are over the thresholds you've set for your species. In our case, we consider a detection a titi duet when the titi templates have a score of 4.0 or higher and titi templates score higher than howler templates.
```{r}
final.results <- template.summary %>% 
  mutate(titi_over_4_and_over_howler = if_else(T >= 4.0 & T > H, 1, 0)
         )
```
